### 2018-01-31 includes/experiment.tex
Removed last sentence of 5.
Removing commented-out sections. Fixing up more content.

\section{Experimental Data and Analysis}
\label{sec:data}

We generated several problem suites with \fuzzer{} that made one
solver perform poorly, but not others. These suites are
\theSuites{}. Figure~\ref{fig:cvc-hard} shows the suites that were
uniquely difficult for \cvc{}. Figure~\ref{fig:z3str3-hard} shows the
suites that were uniquely difficult for \us{}. All experiments were
run in series on the same Linux computer, with a timeout of 15
seconds. \todo{Specify CPU, memory}

@@

We found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \texttt{concats-big} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As an immediate consequence of
testing \us{} on \texttt{concats-big} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly. This allowed us to
move forward with the development of a new heuristic to resolve this
issue and achieve better performance.

%% %VG: this subsection is not needed in the paper
%% \noindent{\textbf{Defects discovered and fixed in \us{}:}} The first
%% issue we found caused \us{} to perform poorly when equivalence classes
%% of arithmetic terms were very large (e.g. the lengths of many strings
%% were equal to 0). This issue was detected by the \texttt{Concats}
%% problem suite. The slowdown was caused by a loop over terms in the
%% same equivalence class to search for an integer constant. The fix was
%% to check the root of the equivalence class for the integer constant,
%% as Z3 makes ``interpreted terms'' (e.g. constants) the root of
%% equivalence classes whenever possible. The second issue, which was
%% exposed by the \texttt{Lengths} suite, caused \us{} to perform poorly
%% when a variable has a large length constraint.  The slowdown was
%% caused by the solver performing a search for a satisfying length even
%% when the input formula explicitly constrained the exact length of a
%% string term.  The fix was to check whether the integer solver had
%% already assigned a length value before searching for length, and skip
%% the search if this was the case.

%% In order to improve the efficiency of any solver, discovering new
%% heuristics is crucial.  As an illustrative example, consider the
%% problem of how the a string solver may detect conflicting
%% constraints. A string solver that gradually propagates
%% (dis)equalities discovered along the way will eventually detect a
%% conflict.  However, in many scenarios, a conflict can be discovered
%% ``early'' before trying to figure out the relationships among
%% subterms and propagating the facts.  The benefits of this eager
%% approach are clear: this saves the effort of reasoning about
%% subterms and propagating observations, as a conflict will
%% eventually be reached in either case, forcing conflict analysis and
%% a backjump.  As reported in existing works, heuristics of this form
%% usually lead to significant improvements.  However, the process of
%% discovering such heuristics is extremely challenging, as they are
%% usually learned from a class of examples whose commonalities are
%% not always obvious or immediately indicative of such
%% inefficiencies.

>>>>

\section{Experimental Results and Analysis}
\label{sec:data}

We generated several problem suites with \fuzzer{} that made one
solver perform poorly, but not others. These suites are
\theSuites{}. Figure~\ref{fig:cvc-hard} shows the suites that were
uniquely difficult for \cvc{}. Figure~\ref{fig:z3str3-hard} shows the
suites that were uniquely difficult for \us{}. All experiments were
run in series, with a timeout of 15 seconds, on the same computer running
Ubuntu Linux 16.04. The computer had 32GB of RAM and an
Intel\textregistered{} Core\texttrademark{} i7-6700 CPU with clock speed
of 3.40GHz.

@@

We found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \textit{concats-big} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As an immediate consequence of
testing \us{} on \textit{concats-big} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly.

### 2018-01-31 includes/introduction.tex
Removing commented-out sections. Fixing up more content.

%VG: add citation to SAT fuzzer
Fuzzing is often used to test all kinds of software,
including SAT solvers. Inspired by their utility, we introduce a
fuzzer for string solvers, \fuzzer{}, and describe its value as an
exploratory validation tool. We demonstrate its efficacy
by presenting limitations it helped us find in
leading string solvers. To the best of our knowledge, \fuzzer{} is the
first tool aimed at automatic generation of string constraints. \fuzzer{} can be
used to mutate or transform existing benchmarks, as well as
randomly generate structured instances. These instances can be scaled with respect
to a variety of parameters, e.g., length of string constants,
depth of concats and regeular expressions, number of variables,
number of length constraints, and many more.

@@

    % Further, StringFuzz's modular design lets users extend it to generate scaling inputs along any parameter of the   input constraint.

\item \textbf{A repository of \smtfull{} problems}:
    We present a repository of \smtfull{} string and regex problem suites we
    generated using \fuzzer{} in Section~\ref{sec:problems}. The repository
    consists of two categories: \texttt{generated}, with new problems generated
    by \fuzzer{}, and \texttt{transformed}, with transformed problems generated
    from a small suite of industrial benchmarks.

>>>>

Fuzzing is often used to test all kinds of software,
including SAT solvers. Inspired by their utility, we introduce a
fuzzer for string solvers, \fuzzer{}, and describe its value as an
exploratory validation tool. We demonstrate its efficacy
by presenting limitations it helped us find in
leading string solvers. To the best of our knowledge, \fuzzer{} is the
first tool aimed at automatic generation of string constraints. \fuzzer{} can
be used to mutate or transform existing benchmarks, as well as
randomly generate structured instances. These instances can be scaled with
respect to a variety of parameters, e.g., length of string constants,
depth of concats and regeular expressions, number of variables,
number of length constraints, and many more.

@@

    \item \textbf{A repository of \smtfull{} problems}:
        We present a repository of \smtfull{} string and regex problem suites
        we generated using \fuzzer{} in section~\ref{sec:problems}. The
        repository consists of two categories: \texttt{generated}, with new
        problems generated by \fuzzer{}, and \texttt{transformed}, with
        transformed problems generated from a small suite of industrial
        benchmarks.

### 2018-01-31 includes/problems.tex
Removing commented-out sections. Fixing up more content.

In this section, we describe the benchmark suites that we generated
with \fuzzer{} on which the experimental evaluation was conducted.

\subsubsection{Generated Benchmarks}

We generated several suites
of string and regex problems using \fuzzer{}.\footnote{The link to the
repository will be added after double-blind review.}
Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.
\begin{table}[t]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Problems have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

\subsubsection{Transformed Benchmarks}

The \texttt{transformed}
suite contains benchmarks derived from existing seed problems by
iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows
the overview. Every problem in the suite is named according to its
seed and the transformations it undertook. For example, input
\texttt{seed-fuzz-graft.smt2} was transformed by applying
\texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.
\begin{table}[t]
    \centering
    \caption{Transformed problem sets generated by \transformer{}.}
    \label{tbl:transformed}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name}   & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
        \textit{Amazon} & Two industrial regex membership problems & 516\\ \hline
        \textit{Regex}  & Z3str2 regular expression test suite     & 10,577\\ \hline
    \end{tabular}
\end{table}

The Amazon category is obtained from two seeds: (1)
\texttt{z3-regex-1.smt2} encodes the assertions \texttt{(S not in
  aaabbb*)} and \texttt{(S in aaa*bbb*)} for a string constant S; (2)
\texttt{z3-regex-2-negative.smt2} includes regex \texttt{(S in
  aaabbb*)} and \texttt{(S not in aaa*bbb*)}.  Both problems come from
our industrial collaborators. In total, 516 unique problems were
generated from these two seeds. The Regex category is seeded by the
z3str2 regular expression test suite \cite{z3str2-tests}.  This test
suite contains 42 problems dealing with a wide variety of regular
expressions.  Through cumulative transformations, we have $10,577$
unique benchmarks.

>>>>

In this section, we describe the benchmark suites that we generated
with \fuzzer{}, and on which we conducted our experimental
evaluation.\footnote{The link to the repository will be added after
double-blind review.}

Table~\ref{tbl:generated} describes the string and regex instances that were
generated by \generator{}. Table~\ref{tbl:transformed} lists benchmarks
derived from existing seed instances by iteratively applying \transformer{}.
Every transformed instance is named according to its
seed and the transformations it undertook. For example, input
\texttt{z3-regex-1-fuzz-graft.smt2} was transformed by applying
\textit{fuzz} and then \textit{graft} to \texttt{z3-regex-1.smt2}.

The \textit{Amazon} category contains 472 instances, derived from two seeds
supplied by our industrial collaborators. The \textit{Regex} category is
seeded by the z3str2 regular expression test suite~\cite{z3str2-tests}. This
test suite contains 42 instances dealing with a wide variety of regular
expressions. Through cumulative transformations, we expanded it to 7,551
instances. Finally, the \textit{Sanitizer} category is obtained from
five industrial e-mail address and IPv4 sanitizers.

\begin{table}[t]
    \centering
    \caption{\generator{}-generated problem sets.}
    \label{tbl:generated}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}             & \textbf{Instances have ...} \\ \hline
        \textit{Concats}          & A single, deep tree of concats. \\ \hline
        \textit{Concats-Extracts} & A single concat tree, and random character extractions. \\ \hline
        \textit{Lengths}          & A single, large length constraint on a variable. \\ \hline
        \textit{Lengths-Concats}  & Variables and their concats, each with fixed lengths. \\ \hline
        \textit{Overlaps}         & A formula of the form A.X = X.B. \\ \hline
        \textit{Regex}            & A test for regex membership of arbitrary complexity. \\ \hline
        \textit{Many-Regexes}     & A test for membership in many random regexes. \\ \hline
        \textit{Regex-Deep}       & A test for membership in a regex with many nested operators. \\ \hline
        \textit{Regex-Pair}       & A test for membership in one regex, but not another. \\ \hline
        \textit{Regex-Lengths}    & A test for membership in a regex, and a length constraint. \\ \hline
    \end{tabular}
\end{table}

\begin{table}[t]
    \centering
    \caption{Transformed problem sets generated by \transformer{}.}
    \label{tbl:transformed}
    \begin{tabular}{|l|l|r|}
        \hline
        \textbf{Name}      & \textbf{Seed}                            & \textbf{Quantity} \\ \hline
        \textit{Amazon}    & Two industrial regex membership problems & 472 \\ \hline
        \textit{Regex}     & Z3str2 regular expression test suite     & 7,551 \\ \hline
        \textit{Sanitizer} & 5 e-mail and IPv4 sanitiser examples     & 1,170 \\ \hline
    \end{tabular}
\end{table}

### 2018-01-31 includes/stringfuzz.tex
Centering and putting tables at the top.
More edits from meeting.
Removing commented-out sections. Fixing up more content.

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
problems. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        Table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        Table~\ref{tbl:transformers}. Note that transformers
        \textit{translate} and \textit{reverse} also preserve problem
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[H]
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{concats}     & Long concats and optional random extracts. \\ \hline
        \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{regex}       & Regexes of varying complexity. \\ \hline
        \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{nop}         & Does nothing. \\ \hline
        \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

\fuzzer{} can also transform the instances it generates. All of the
transformers in \transformer{} can operate on regex terms. They can
multiply and permute regex literals (e.g. \texttt{reverse},
\texttt{multiply}, \texttt{translate}), randomly replace regex
operators (e.g. \texttt{fuzz}), and manipulate the syntax tree
(e.g. \texttt{graft}, \texttt{rotate}).

@@

\fuzzer{} can be used to
generate benchmark suites in a batch mode, or can be used in an online
randomized fashion for exploratory debugging. For example, \fuzzer{}
can be invoked as follows to feed random problems to \cvc{} until it
returns an error:
% \vspace{-0.15in}
% \begin{figure}[H]
%     \centering
%     \begin{minipage}[b]{0.91\linewidth}
%         \begin{minted}{bash}
{\scriptsize\begin{verbatim}
while \
  stringfuzzg -r random-ast -m \
  | tee problem.smt25 \
  | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}
%         \end{minted}
%     \end{minipage}
% \end{figure}
% \vspace{-0.15in}

>>>>

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
problems. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} problems. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        problems. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{translate} and \textit{reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        problem as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

\begin{table}[t]
    \centering
    \caption{\generator{} generators.}
    \label{tbl:generators}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
        \textit{concats}     & Long concats and optional random extracts. \\ \hline
        \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
        \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
        \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
        \textit{regex}       & Regexes of varying complexity. \\ \hline
        \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
        \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
    \caption{\transformer{} transformers.}
    \label{tbl:transformers}
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Name}        & \textbf{The transformer ...} \\ \hline
        \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
        \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
        \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
        \textit{nop}         & Does nothing. \\ \hline
        \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
        \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
        \textit{translate}   & Maps every character X to a different character Y. \\ \hline
        \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
    \end{tabular}
\end{table}

@@

\fuzzer{} can also transform the instances it generates. All of the
transformers in \transformer{} can operate on regex terms. They can
multiply and permute regex literals (e.g. \textit{reverse},
\textit{multiply}, \textit{translate}), randomly replace regex
operators (e.g. \textit{fuzz}), and manipulate the syntax tree
(e.g. \textit{graft}, \textit{rotate}).

@@

\fuzzer{} can be used to
generate benchmark suites in a batch mode, or can be used in an online
randomized fashion for exploratory debugging. For example, \fuzzer{}
can be invoked as follows to feed random problems to \cvc{} until it
returns an error:
% \vspace{-0.15in}
% \begin{figure}[H]
%     \centering
%     \begin{minipage}[b]{0.91\linewidth}
%         \begin{minted}{bash}
{\scriptsize\begin{verbatim}
while stringfuzzg -r random-ast -m | tee problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}
%         \end{minted}
%     \end{minipage}
% \end{figure}
% \vspace{-0.15in}
